\pkg{cleanR}: Your Personal Maid for Cleaning Datasets in \proglang{R}

### 2016-10-11 latex/article.Rnw
Fixed paper

Data cleaning is a time consuming endeavour and it inherently requires
human interaction since every dataset is different and the variables
in the dataset can only be understood in the proper context of the
experiment. While each dataset is different and require
... (``simple questions''

... there are a number of common occurrences that .... This is particulary
Automated

... partly be done by ... However ... enough ... similar tasks that
pop

However it is not uncommon for ... to run a set of ....

In this paper, we will briefly present some properties of the simplex
distribution, which provides a foundation for the inference and
computation in the package simplexreg (Zhang, Qiu, and Shi 2016). The
package is implemented in the R system and available from the
Comprehensive R Archive Network (CRAN) at https:
//CRAN.R-project.org/package=simplexreg. There are three major lines
of functions:
1. Calculation of the density, cumulative distribution, quantile and
random number gen- erating functions for the simplex distribution.
2. Statistical inference in the simplex generalized linear model
(SGLM, Zhang and Qiu 2014) via maximum likelihood (ML) for
cross-sectional proportional data set.
3. Analysis of longitudinal proportional data set via an extended
version of generalized estimating equations (GEE) using the simplex
distribution.

@@

\section{The structure of \pkg{cleanR}} \label{sec:internals}

\section{Extending \pkg{cleanR} by adding custom error checks} \label{sec:extending}

>>>>

Data cleaning is a time consuming endeavour and it inherently requires
human interaction since every dataset is different and the variables
in the dataset can only be understood in the proper context of the
experiment. While each dataset is different and requires unique
attention there are often a number of similar tasks that are
undertaken as part of the data cleaning and quality control
process. This is especially true when data are received repeatedly
from the same source as data providers tend to use a stable setup (and
hence introduce the same set of potential mistakes that need to be
identified and corrected).

In many situations these errors are discovered in the process of the
data analysis (e.g., a categorical variable with numeric labels is
wrongly classified as a numeric variable), but in other cases a human
with knowledge about the data context area is needed to identify
possible mistakes in the data (e.g., if there are 4 categories for a
variable that should only have 3).  It is necessary to summarize
information --- both numerically and graphically --- about each
variable in order for an investigator to detect possible errors, and
help raise flags of warning to draw the attention of investigator for
the situations where there \emph{might} be a error.

The \pkg{cleanR} package supports the automated checking of errors in
a dataset and produces an output document with detailed information
about each variable that can be
The process is illustrated in Figure~\ref{fig:flowchart1}. A dataset
is cleaned and an \proglang{R} markdown file is produced and possibly
rendered into html or pdf.

@@

\section{The structure of cleanR} \label{sec:internals}

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm]
\node (start) [io] {data.frame};
\node (pro1) [process, below of=start] {prechecks};
\node (screen) [io, right of=pro1, xshift=2cm] {Screen};
\node (pro2) [process, below of=pro1] {Rmd};
\node (html) [io, right of=pro2, xshift=2cm] {html};
\node (pdf) [io, below of=html] {pdf};
\draw [arrow] (start) -- (pro1);
\draw [arrow] (pro1) -- (screen);
\draw [arrow] (pro1) -- (pro2);
\draw [arrow] (pro2) -- (html);
\draw [arrow] (pro2) -- (pdf);
\end{tikzpicture}
\end{center}
\label{fig:flowchart1}
\caption{The process of cleaning data frames using the \pkg{cleanR}
  package. A series of check is applied to each variable in a
  \texttt{data.frame} and a summary of the result is either printed to
  the screen, or an \proglang{R} markdown file is produced which is
  subsequently rendered.}
\end{figure}

Bør vise DF -> prechecks -> for each variable do this ...

\section{Extending cleanR by adding custom error checks} \label{sec:extending}

### 2016-10-11 latex/article.tex
Fixed paper

\section{Introduction}
Statisticians and data analysts spend a large portion of their time on
data cleaning and on data wrangling. Packages such as \pkg{data.table}, \pkg{plyr},
and \pkg{} make data wrangling a lot easier in R, but there are only a
few options available for automated data cleaning.

Data cleaning is a time consuming endeavour and it inherently requires
human interaction since every dataset is different and the variables
in the dataset can only be understood in the proper context of the
experiment. While each dataset is different and require
... (``simple questions''

... there are a number of common occurrences that .... This is particulary
Automated

... partly be done by ... However ... enough ... similar tasks that
pop

However it is not uncommon for ... to run a set of ....

In this paper, we will briefly present some properties of the simplex
distribution, which provides a foundation for the inference and
computation in the package simplexreg (Zhang, Qiu, and Shi 2016). The
package is implemented in the R system and available from the
Comprehensive R Archive Network (CRAN) at https:
//CRAN.R-project.org/package=simplexreg. There are three major lines
of functions:
1. Calculation of the density, cumulative distribution, quantile and
random number gen- erating functions for the simplex distribution.
2. Statistical inference in the simplex generalized linear model
(SGLM, Zhang and Qiu 2014) via maximum likelihood (ML) for
cross-sectional proportional data set.
3. Analysis of longitudinal proportional data set via an extended
version of generalized estimating equations (GEE) using the simplex
distribution.

>>>>

\section{Introduction}
Statisticians and data analysts spend a large portion of their time on
data cleaning and on data wrangling. Packages such as
\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot easier in
\proglang{R}, but there are only a few options available for automated data
cleaning.

Data cleaning is a time consuming endeavour and it inherently requires
human interaction since every dataset is different and the variables
in the dataset can only be understood in the proper context of the
experiment. While each dataset is different and requires unique
attention there are often a number of similar tasks that are
undertaken as part of the data cleaning and quality control
process. This is especially true when data are received repeatedly
from the same source as data providers tend to use a stable setup (and
hence introduce the same set of potential mistakes that need to be
identified and corrected).

In many situations these errors are discovered in the process of the
data analysis (e.g., a categorical variable with numeric labels is
wrongly classified as a numeric variable), but in other cases a human
with knowledge about the data context area is needed to identify
possible mistakes in the data (e.g., if there are 4 categories for a
variable that should only have 3).  It is necessary to summarize
information --- both numerically and graphically --- about each
variable in order for an investigator to detect possible errors, and
help raise flags of warning to draw the attention of investigator for
the situations where there \emph{might} be a error.

The \pkg{cleanR} package supports the automated checking of errors in
a dataset and produces an output document with detailed information
about each variable that can be
The process is illustrated in Figure~\ref{fig:flowchart1}. A dataset
is cleaned and an \proglang{R} markdown file is produced and possibly
rendered into html or pdf.
